---
title: "Exploring the data"
author: "José Alberto Arcos Sánchez"
date: "24 de julio de 2017"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(jsonlite)
library(purrr)
library(lubridate)
library(glue)
```


### Fast exploration


Let's load a json file and see what it contains:

```{r}
file <- read_json("../2017/07/21/405540905874_CloudTrail_eu-west-1_20170721T0000Z_toFMCcKusnum4iUt.json")
str(file)
```

```{r}
map(file[[1]], ~ list(length(.), names(.)))
```

Each element in the json file contains 14 or 15 features. The "extra" feature is *apiVersion*. Let's see if the existence of apiVersion is related to other variable:

```{r}
res <- map(file[[1]], ~ list("apiVersion" %in% names(.), .$eventSource))
do.call(rbind, res)
```

It looks like the existence of *apiVersion* depends on the source of the event. Nice!

To facilitate the exploration, let's create a dataframe with the json file.

The data contains some complexities:
- *userIdentity* contains a list with varying elements. We need to take its elements out:

```{r}
map(file[[1]], ~ names(.$userIdentity)) %>% unlist() %>% unique()
```

- *sessionContext* is also a list with varying elements:

```{r}
map(file[[1]], ~ names(.$userIdentity$sessionContext)) %>% unlist() %>% unique()
map(file[[1]], ~ names(.$userIdentity$sessionContext$attributes)) %>% 
  unlist() %>% unique()
map(file[[1]], ~ names(.$userIdentity$sessionContext$sessionIssuer)) %>% 
  unlist() %>% unique()
```

This may happen more times in our unseen data, so let's look for a general solution. We can build a data frame unnesting every value using the **plyr** package:

```{r}
library(plyr)

data <- file[[1]]
datalist <- map(data, ~ as.data.frame(as.list(unlist(.))))
df <- rbind.fill(datalist)
View(df)
```

Great!! 

--------------------------------------------------------------------------------

Some files contain fields that are difficult to "tidy":

```{r}
file <- read_json("../2017/04/22/405540905874_CloudTrail_eu-west-1_20170422T0000Z_EdBbEouphu5PwnGH.json")
data <- file[[1]]
datalist <- map(data, ~ as.data.frame(as.list(unlist(.))))
df <- rbind.fill(datalist)
View(df)
```

*df* has 596 variables. Why!!??

```{r}
colnames(df)
```

Because of the *responseElements* component. Let's see the tree structure:

```{r}
str(data[[254]])
```


### Exploration by variable (only main variables)

Let's take data for an entire day.

```{r}
json_files <- list.files(path = "../2017/07/21/", full.names = TRUE)
json_list <- map(json_files, ~ read_json(path = .)[[1]])
json_list <- do.call(c, json_list)
```

And see how many events we have and what they contain:

```{r}
glue("Number of events: {length(json_list)}\n\n")
glue("Number of fields in each event:")
table(map_dbl(json_list, ~ length(.)))
```

The number of fields in each event is not constant. Some events have more fields than others.
Let's see all the possible fields:

```{r}
map(json_list, ~ names(.)) %>% unlist() %>% table() %>% sort(decreasing = TRUE)
```

20 possible fields! Let's see how many different combinations are seen in a day:

```{r}
all_combinations <- map(json_list, ~ names(.))
all_unique_combinations <- unique(all_combinations)
glue("Number of different combinations = {length(all_unique_combinations)}")
```

Maybe there are 6 types of events??


Let's now explore the main fields and their typical values:

--------------------------------------------------------------------------------
**awsRegion**

```{r}
table(map_chr(json_list, ~ .$awsRegion))
```


--------------------------------------------------------------------------------
**eventID**

```{r}
map_chr(json_list, ~ .$eventID) %>% unique() %>% head()
```


--------------------------------------------------------------------------------
**eventName**

This looks like an important field. It may describe the action the "user" wants to perform. For this particular day, 87 different event names have been logged.

```{r}
glue("Number of different names: {map_chr(json_list, ~ .$eventName) %>% unique() %>% length()}")
sort(table(map_chr(json_list, ~ .$eventName)), decreasing = TRUE)
```


--------------------------------------------------------------------------------
**eventSource**

There are 36 different sources in this particular day.

```{r}
sort(table(map_chr(json_list, ~ .$eventSource)), decreasing = TRUE)
```

The most frequent one in *ec2.amazonaws.com* followed by *dynamodb.amazonaws.com* and *s3.amazonaws.com*.


--------------------------------------------------------------------------------
**eventTime**

```{r}
datetimes <- map(json_list, ~ ymd_hms(.$eventTime)) %>% do.call(c,.)
summary(datetimes)
```

It contains events of a unique day, as expected. It is not a complete day, so the results are nice.


--------------------------------------------------------------------------------
**eventType**

```{r}
sort(table(map_chr(json_list, ~ .$eventType)), decreasing = TRUE)
```


--------------------------------------------------------------------------------
**eventVersion**

There are at least two event versions:

```{r}
table(map_chr(json_list, ~ .$eventVersion))
```

The 1.05 version is much more frequent.


--------------------------------------------------------------------------------
**recipientAccountId**

```{r}
table(map_chr(json_list, ~ .$recipientAccountId))
```


--------------------------------------------------------------------------------
**requestParameters**

A list with lots of fields:

```{r}
map(json_list, ~ names(.$requestParameters)) %>% unlist() %>% table() %>% sort(decreasing = TRUE) %>% head(n = 10)
```

--------------------------------------------------------------------------------
**responseElements**

A list with lots of fields:

```{r}
map(json_list, ~ names(.$responseElements)) %>% unlist() %>% table() %>% sort(decreasing = TRUE) %>% head(n = 10)
```


--------------------------------------------------------------------------------
**sourceIPAddress**

There are not a lot of source IP addresses in this particular day.

```{r}
sort(table(map_chr(json_list, ~ .$sourceIPAddress)), decreasing = TRUE) %>% head()
```


--------------------------------------------------------------------------------
**userAgent**

```{r}
sort(table(map_chr(json_list, ~ .$userAgent)), decreasing = TRUE) %>% head()
```

--------------------------------------------------------------------------------
**userIdentity**

This field is a list with several (and changing) subfields:

```{r}
sort(table(map(json_list, ~ names(.x$userIdentity)) %>% unlist()), decreasing = TRUE)
```


*type*

```{r}
sort(table(map_chr(json_list, ~ .x$userIdentity$type)), decreasing = TRUE)
```


*accountId*

```{r}
sort(table(map(json_list, ~ .x$userIdentity$accountId) %>% unlist()), decreasing = TRUE)
```


*principalId*

```{r}
sort(table(map(json_list, ~ .x$userIdentity$accountId) %>% unlist()), decreasing = TRUE)
```


*arn*

```{r}
sort(table(map(json_list, ~ .x$userIdentity$arn) %>% unlist()), decreasing = TRUE)
```


*accessKeyId*

```{r}
sort(table(map(json_list, ~ .x$userIdentity$accessKeyId) %>% unlist()), decreasing = TRUE)
```


*sessionContext*

It is a list with 2 fields:

```{r}
sort(table(map(json_list, ~ names(.x$userIdentity$sessionContext)) %>% unlist()), decreasing = TRUE)
```

*sessionContext.attributes*

```{r}
sort(table(map(json_list, ~ names(.x$userIdentity$sessionContext$attributes)) %>% unlist()), decreasing = TRUE)
```

mfaAuthenticated >> (TRUE/FALSE)
creationDate >> date :)

*sessionContext.sessionIssuer*

```{r}
sort(table(map(json_list, ~ names(.x$userIdentity$sessionContext$sessionIssuer)) %>% unlist()), decreasing = TRUE)
```


*userName*

```{r}
sort(table(map(json_list, ~ .x$userIdentity$userName) %>% unlist()), decreasing = TRUE)
```


*invokedBy*

```{r}
sort(table(map(json_list, ~ .x$userIdentity$invokedBy) %>% unlist()), decreasing = TRUE)
```








--------------------------------------------------------------------------------
**requestID**

```{r}
map(json_list, ~ .$requestID) %>% unlist() %>% unique() %>% head()
```


--------------------------------------------------------------------------------
**apiVersion**

```{r}
sort(table(map_chr(json_list, ~ ifelse(test = is.null(.x$apiVersion), yes = "unknown", no = .x$apiVersion))), decreasing = TRUE)
```

--------------------------------------------------------------------------------
**errorCode**

```{r}
map(json_list, ~ .$errorCode) %>% unlist() %>% table() %>% sort(decreasing = TRUE)
```

--------------------------------------------------------------------------------
**errorMessage**

```{r}
map(json_list, ~ .$errorMessage) %>% unlist() %>% table() %>% sort(decreasing = TRUE)
```

--------------------------------------------------------------------------------
**resources**

A list with three fields:

```{r}
map(json_list, ~ names(.$resources %>% unlist(recursive = FALSE))) %>% unlist() %>% table()
```


*resources.accountId*

```{r}
map(json_list, ~ .$resources[[1]]$accountId) %>% unlist() %>% table()
```


*resources.ARN*

```{r}
map(json_list, ~ .$resources[[1]]$ARN) %>% unlist() %>% table() %>% head()
```


*resources.type*

```{r}
map(json_list, ~ .$resources[[1]]$type) %>% unlist() %>% table() %>% head()
```

--------------------------------------------------------------------------------
**sharedEventID**

```{r}
map(json_list, ~ .$sharedEventID) %>% unlist() %>% table() %>% head()
```


--------------------------------------------------------------------------------
**additionalEventData**

A list with several varying fields:

```{r}
map(json_list, ~ names(.$additionalEventData)) %>% unlist() %>% table()
```