---
title: "Redshift cluster usage"
author: "José Alberto Arcos Sánchez"
date: "28 de julio de 2017"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(purrr)
library(futile.logger)
library(glue)
library(stringr)
library(jsonlite)
library(tidyverse)
library(lubridate)
library(magrittr)
```

### Abstract

This document explores the usage of Redshift clusters in the available log data. Redshift clusters are one of the main economical costs of the project, so it is important to monitor them in order to detect inefficient patterns (like clusters running but doing nothing).

<br>

### Redshift events in Cloudtrail

The list of supported actions and its meaning can be found [here](http://docs.aws.amazon.com/redshift/latest/APIReference/API_Operations.html).

We can easily take all the events whose *eventSource* is *redshift.amazonaws.com*:

```{r}

folders <- list.dirs("../")

events <- map(folders, function(folder){
  files <- list.files(path = folder, full.names = T)
  files <- files[str_detect(files, ".json")]
  map(files, function(file){
    flog.debug(glue("Reading file {file}"))
    events_in_file <- read_json(file) %>% unlist(recursive = FALSE)
    map(events_in_file, function(event){
      if(event$eventSource == "redshift.amazonaws.com"){
        return(event)
      } else {
        return(NULL)
      }
    })
  })
})

```

The events variable is huge (5.6 Gb), so let's clean it and save it to avoid possible accidents:

```{r}
events <- unlist(events, recursive = FALSE) %>% unlist(recursive = FALSE)
is_null <- map_lgl(events, ~ is.null(.))
events <- events[!is_null]
save(events, file = "all_redshift_events.RData")
```

Let's see what kinds of events exist in out dataset:

```{r}
map_chr(events, ~ .$eventName) %>% table() %>% sort(decreasing = TRUE)
```

<br>

### Creation and deletion events (I)

The events that may interest us are:

[DeleteCluster](http://docs.aws.amazon.com/redshift/latest/APIReference/API_DeleteCluster.html)
[RestoreFromClusterSnapshot](http://docs.aws.amazon.com/redshift/latest/APIReference/API_RestoreFromClusterSnapshot.html)

So lets take only the events of interest:

```{r}
is_interesting <- map_lgl(events, ~ .$eventName %in% c("DeleteCluster", "RestoreFromClusterSnapshot"))
events <- events[is_interesting]
save(events, file = "interesting_events.RData")
```

Now, let's plot a timeline with all the interesting events. We need to create a dataframe with dates and eventNames:

```{r}
load("interesting_events.RData")
plot_datetimes <- map_chr(events, ~ .$eventTime) %>% ymd_hms()
plot_events <- map_chr(events, ~ .$eventName)
df_timeline <- tibble(datetime = plot_datetimes, event = plot_events)

ggplot(data = df_timeline %>% filter(event != "DescribeReservedNodes"), 
       mapping = aes(x = datetime, y = event)) + 
  geom_point(mapping = aes(shape = event, color = event)) +
  guides(color = FALSE, shape = FALSE) +
  theme_minimal()
```

Wow, lots of events. Let's see a summary:

```{r}
table(df_timeline$event)
```

Well, something extrange here. Clusters have been destroyed 198 times, while created only 106 times... 

What is happening?

It would be nice if we could see the exact number of clusters in each moment. Is that possible? Let's explore the *DescribeClusters* events...

<br>

### Exploring the *DescribeClusters* events

What about the *DescribeClusters* events?

```{r}
load("all_redshift_events.RData")
describe_clusters_index <- map_lgl(events, ~ .$eventName == "DescribeClusters")
describe_clusters <- events[describe_clusters_index]
rm(events)
```

Yeah! This event contains info about the existing clusters, and we have lots of this kind of events! 
Let's create a dataframe containing all the interesting cluster info. We want to know all the clusters that have existed, when they were created and the number of nodes of each cluster:

```{r}
# Preallocate variable
df_datetime = rep(NA_character_, 5e5)
df_cluster = rep(NA_character_, 5e5)
df_creation_date = rep(NA_character_, 5e5)
df_n_nodes = numeric(5e5)

# Fill variable
index <- 1
pb <- progress_estimated(n = length(describe_clusters))
# pb <- progress_estimated(n = 1e4)
for(i in 1:length(describe_clusters)){
  # for(i in 1:1e4){  
  event <- describe_clusters[[i]]
  n_clusters <- length(event$responseElements$clusters)
  
  if(n_clusters > 0){
    
    datetimes <- rep(event$eventTime, n_clusters)
    cluster_names <- map_chr(event$responseElements$clusters,  ~ .$clusterIdentifier)
    creation_dates <- map_chr(event$responseElements$clusters, function(cluster){
      times <- cluster$clusterCreateTime
      ifelse(is.null(times), NA_character_, times)
    })
    n_nodes <- map_int(event$responseElements$clusters, ~ .$numberOfNodes)
    
    if(length(datetimes) != n_clusters) browser()
    if(length(cluster_names) != n_clusters) browser()
    if(length(creation_dates) != n_clusters) browser()
    if(length(n_nodes) != n_clusters) browser()
    
    df_datetime[index:(index + n_clusters - 1)] <- datetimes
    df_cluster[index:(index + n_clusters - 1)] <- cluster_names
    df_creation_date[index:(index + n_clusters - 1)] <- creation_dates
    df_n_nodes[index:(index + n_clusters - 1)] <- n_nodes
    
    index <- index + n_clusters
    
  }

  pb$tick()$print()
}

cluster_info <- tibble(datetime = df_datetime[1:index],
                       cluster = df_cluster[1:index],
                       creation_date = df_creation_date[1:index],
                       n_nodes = df_n_nodes[1:index])

cluster_summary <- 
  cluster_info %>% filter(complete.cases(.)) %>% 
  group_by(cluster, creation_date, n_nodes) %>% 
  summarise(datetime = datetime[[1]]) %>% 
  ungroup() %>% 
  mutate(datetime = ymd_hms(datetime)) %>% 
  mutate(creation_date = fast_strptime(creation_date, format = "%b %d, %Y %H:%M:%S %p", lt = FALSE))

cluster_summary
```

Let's plot the creation of each cluster:

```{r}
ggplot(data = cluster_summary, mapping = aes(x = creation_date, y = cluster)) +
  geom_point(mapping = aes(size = n_nodes), color = "blue", alpha = 0.3) +
  guides(size = FALSE) +
  theme_minimal() +
  theme(axis.title = element_blank())
```

Nice plot!! Each point represents a cluster creation. Its size represents the number of nodes of the cluster (from 1 to 10).
There are clear patterns in the use of most of the clusters. Nice!

Let's see if we can cross this data with the creation and deletion events.

<br>

### Creation and deletion events (II)

When a cluster is deleted, the deletion event (*DeleteCluster*) contains the identifier of the cluster (*$requestParameters$clusterIdentifier*). In the same way, when a cluster is created, the creation event (*RestoreFromClusterSnapshot*) contains the cluster identifier (*$responseElements$clusterIdentifier*).

Let's create a dataframe containing all the creation and deletion events with their date and their target cluster:

```{r}
load("interesting_events.RData")

n_events <- length(events)
df_event_names    <- rep(NA_character_, n_events)
df_event_times    <- rep(NA_character_, n_events)
df_target_cluster <- rep(NA_character_, n_events)

for(i in 1:n_events){
  event <- events[[i]]
  df_event_names[i] <- event$eventName
  df_event_times[i] <- event$eventTime
  
  cluster_id <- event$responseElements$clusterIdentifier
  df_target_cluster[i] <- ifelse(is.null(cluster_id), NA_character_, cluster_id)
}

creation_deletion_events <- tibble(event = df_event_names,
                                   datetime = df_event_times %>% ymd_hms(),
                                   target_cluster = df_target_cluster)

```

Let's plot the events:

```{r}

ggplot(data = creation_deletion_events, 
       mapping = aes(x = datetime, y = target_cluster)) +
  geom_point(mapping = aes(color = event)) +
  theme_bw() +
  theme(legend.position = "bottom", 
        axis.title.x = element_blank(),
        axis.title.y = element_blank(),
        legend.title = element_blank())

```

It looks that the events are correct! The problem is that lots of them do not have a target. Interesting detail.
